\name: Combining Bottom-up and top-down approaches through graph learning over interaction networks for drug-target-interaction prediction

### 2021-03-21 paper/main.tex
.

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. Molecular information can be used
to directly determine whether two molecules have the ability to
interact,

What features localize on graph? Molecular interactions based on
structure not likely -- only possible to find binding sites to other
proteins.

Phenotype-based features localize...

\todo[inline]{Methods:}
Through the very nature of the graph convolutional neural network, we
build the transformed representation for all proteins in every
forwarding step of the model. Note particularly, that the same
convolutional filter and weight matrix are applied to and learned for
all nodes simultaneously. By construction, for a single drug we can
compute and predict all its interactors in a single run of the model,
leading to significantly less computing time.

using indications is like cheating, as not applicable for searching new drugs

drug indications are highly predictive for downstream tasks, but lack capability to differentiate highly related drugs/proteins

@@

protein functions localize on the PPI graph, while molecular features don't

@@

>>>>

Another view on these two approaches is as direct and indirect ways to
predict drug--target interactions. Molecular information can be used
to directly determine whether two molecules (such as a drug and
protein) have the ability to interact, whereas information about
phenotypic consequences of a drug (drug effects) or disruption of a
protein function can be used to indirectly suggest candidate
drug--target interactions. Molecular information will be specific to a
drug--target pair and we would not expect this information to
propagate through a protein--protein interaction network; the main
information about drug--target interactions that could be obtained
from interactions between proteins is information about binding sites
between proteins that may also be used by a drug molecule (i.e.,
information that protein $P_1$ binds to protein $P_2$ reveals
information about the molecular structures of both $P_1$ and
$P_2$). On the other hand, phenotypic consequences of changes in
protein function, or drug effects, are often a result of aberrant
pathway or network activity and involve more than one protein;
consequently, we expect these features to benefit more from including
information about protein--protein interactions. Our results confirm
this hypothesis and demonstrate that molecular features do not benefit
from including the interaction network whereas the indirect, top-down
features benefit from the propagation over the interaction network.

% \todo[inline]{Methods:}
% Through the very nature of the graph convolutional neural network, we
% build the transformed representation for all proteins in every
% forwarding step of the model. Note particularly, that the same
% convolutional filter and weight matrix are applied to and learned for
% all nodes simultaneously. By construction, for a single drug we can
% compute and predict all its interactors in a single run of the model,
% leading to significantly less computing time.

There are other types of indirect features that could be added to our
model. A common feature that is added are drug indications which are
highly predictive of drug--target interactions \cite{}; however, we do
not include them in our model as including drug indications would
allow our model to make many trivial predictions based only on
remembering which targets are often used for which indication;
including network information would likely benefit predictions based
on drug indications because different drugs may target the same
pathway through different mechanisms \cite{}.

@@

@@

We spent a significant amount of time testing different evaluation
schemes for predicting drug--target interactions. This was motivated
by our finding that a na\"ive predictor can achieve a performance only
slightly below state of the art drug--target prediction methods.
